Optimal updating factor in adaptive flat-distribution-sampling simulations

### 2017-01-24 doc/invt.tex
invt.tex: cleaning up Appendix F

First, we need to define the error in the bias potential.
%
This is done by subtracting from the bias potential
two contributions.
%
The first contribution comes from the target value,
Eq. \eqref{eq:Vi_target}.
%
For this purpose,
we introduce a shifted bias potential
%
\begin{equation}
  v_i(t)
  \equiv
  V_i(t)
  -
  \ln \frac { p^*_i }
            { p_i }
  ,
  \label{eq:v_def}
\end{equation}
%
which should tend to a constant of $i$
upon convergence.
Note that the updates to $v_i$'s are
the same as the updates to $V_i$'s,
since the shift
remains a constant during the course of simulation,
%%
%In terms of $v_i$'s, Eq. \eqref{eq:pi_p_V}
%becomes
%%
%\begin{equation}
%  \pi_i
%  =
%  \frac{                p_i \, e^{-v_i} }
%       { \sum_{j = 1}^n p_j \, e^{-v_j} }
%  \propto
%  p_i \, e^{-v_i}.
%  \label{eq:pi_p_v}
%\end{equation}
%%
%According to Eq. \eqref{eq:pi_p_V1},

Second, the bias potential following Eq. \eqref{eq:wl_update}
generally grows over time.
%
Since a uniform increase of $V_i(t)$, or, equivalently, that of $v_i(t)$,
does not affect the
resulting distribution $\pi_i(t)$,
%
the real deviation is more readily defined
from the difference between $v_i(t)$ and a baseline value, $\bar v(t)$,
that represents the average growth\cite{
dama2014},
%
\begin{equation}
  x_i(t) \equiv v_i(t) - \bar v(t)
  %=
  %v_i - \sum_{j = 1}^n \rho_j \, v_j
  ,
\label{eq:x_def}
\end{equation}
%
where
%
\begin{equation}
  \bar v(t)
  = \sum_{i = 1}^n \rho_i \, v_i(t)
  ,
\label{eq:vbar_def}
\end{equation}
for some distribution,
$\pmb \rho = (\rho_1, \dots, \rho_n)$.
%
In this way,
the variables, $x_i(t)$'s,
represent the error of the bias potential,
and are expected to vanish on convergence.
%
We, therefore, define
the error of the bias potential as
%
\begin{equation}
  \Err
  =
  \sum_{i = 1}^n \rho_i \, \left\langle x_i^2(t) \right\rangle
  .
\label{eq:error_sum}
\end{equation}

@@

To proceed, we will
approximate the finite difference Eq. \eqref{eq:wl_update}
by a differential equation
%
\begin{equation}
  \dot V_i(t)
  \equiv
  \frac{ dV_i(t) } { dt }
  =
  \alpha(t) \, \frac{ h_i(t) } { p_i },
  \label{eq:vt_diffeq}
\end{equation}
%
with
%
$
h_i(t) = \delta_{i, i(t)}
$
%\begin{equation}
%  h_i(t) = \delta_{i, i(t)}
%  ,
%  \label{eq:h_def}
%\end{equation}
%
being the instantaneous histogram,
which is equal to $1.0$
for the bin $i(t)$ at time $t$
or zero otherwise.

Next, we split $h_i(t)$ into a deterministic averaged part
and a stochastic part, $\zeta_i(t)$,
%
\begin{equation}
  h_i(t) = \langle h_i(t) \rangle + \zeta_i(t).
  \label{eq:h_split}
\end{equation}
%
The deterministic part can be related
to the ``ensemble average'' of $h_i$,
where the ensemble consists of many similar simulation copies
sharing the same schedule $\alpha(t)$
the same bias potential at time $t$.
%
The initial states and the stochastic forces
during the process may differ, however.

@@

The average growth rate of $v(t)$ is given by
\begin{equation}
  \frac{ d \bar v (t) }
       { d t }
  =
  \alpha(t)
  \sum_{ r = 1 }^n \rho_r \, \frac{ h_r(t) } { p_r }
  .
\label{eq:dvbardt}
\end{equation}
%

From Eqs.
\eqref{eq:x_def},
\eqref{eq:rhox_sum}-\eqref{eq:dvbardt},
%\eqref{eq:rhox_sum},
%\eqref{eq:vt_diffeq},
%\eqref{eq:h_split}
%\eqref{eq:h_ave},
%\eqref{eq:dvbardt},
we get a set of decoupled equations
%
\begin{equation}
  \dot x_i(t)
  =
  -\alpha(t) \, \left[ x_i(t) - \xi_i(t) \right]
  ,
\label{eq:dxdt_WL}
\end{equation}
where
\begin{equation}
  \xi_i(t)
  =
  \frac{ \zeta_i(t) }
       { p_i }
  -
  \sum_{ r = 1 }^n
    \rho_r
    \frac{ \zeta_r(t) }
         {     p_r    }
  .
\label{eq:xi_def}
\end{equation}
\note{The derivation:
$$
\begin{aligned}
  \dot x_i
  &\stackrel{\mathrm{Eq.\; \eqref{eq:x_def}}}
            {=\joinrel=\joinrel=\joinrel=\joinrel=}
  \dot v_i - \dot{ \bar v }
  \stackrel{\mathrm{Eqs.\; \eqref{eq:vt_diffeq}\; and\; \eqref{eq:dvbardt}}}
            {=\joinrel=\joinrel=\joinrel=\joinrel=\joinrel=\joinrel=\joinrel=\joinrel=\joinrel=\joinrel=}
  \alpha(t) \left(
    \frac{ h_i } { p_i }
    -
    \sum_{ r = 1 }^n
      \rho_r \frac{ h_r } { p_r }
  \right)
  \\
  &\stackrel{\mathrm{Eq.\; \eqref{eq:h_split}}}
            {=\joinrel=\joinrel=\joinrel=\joinrel=}
  \alpha(t) \left(
    \frac{ \langle h_i \rangle } { p_i }
    -
    \sum_{ r = 1 }^n
      \rho_r \frac{ \langle h_r \rangle } { p_r }
    +
    \xi_i(t)
  \right)
  \\
  &\stackrel{\mathrm{Eqs.\; \eqref{eq:rhox_sum} \;and\; \eqref{eq:h_ave}}}
            {=\joinrel=\joinrel=\joinrel=\joinrel=\joinrel=\joinrel=\joinrel=\joinrel=\joinrel=\joinrel=\joinrel=}
  \alpha(t) \left[ -x_i(t) + \xi_i(t) \right].
\end{aligned}
$$
}

The total error defined in Eq. \eqref{eq:error_sum}
now depends implicitly on the function $\alpha(t)$
via $x_i(t)$.
%
Our aim is to find the $\alpha(t)$
that minimizes the error
at the end of the simulation period $T$.
%
It suffices to consider
a one-variable problem that
minimizes each individual $\left\langle x_i^2 \right\rangle$.

@@

The error function defined in Eq. \eqref{eq:error_sum}
can be rewritten in terms of the eigenmodes of $\mathbf w$
\begin{align}
  \Err
  \equiv
  \sum_{i = 1}^n \rho_i \,
    \left\langle x_i^2 \right\rangle
  =
  \sum_{i, j, k=1}^n \phi_{ki} \, \phi_{kj} \,
    \left\langle x_i \, x_j \right\rangle
  =
  \sum_{k = 1}^n
    \left\langle
      y_k^2
    \right\rangle
  ,
\notag
\end{align}
%
where we have used Eq. \eqref{eq:eig_orthonormal_cols}.
%
From Eq. \eqref{eq:eigenmode1},
the first eigenmode,
$y_1 = \sum_{j=1}^n \rho_j \, x_j$,
vanishes identically by Eq. \eqref{eq:rhox_sum}.
%
Thus, we can start the last sum %in Eq. \eqref{eq:y2_sum}
from $k = 2$,
\begin{equation}
  \Err
  =
  \sum_{k = 2}^n
    \left\langle
      y_k^2
    \right\rangle
  .
\label{eq:y2_sum}
\end{equation}

@@

We have also shown that
the single-bin scheme is optimal
in the long time limit
without a priori assumptions
about the smoothness of the PMF.
%
More generally, with perfect filtering
of certain eigenmodes, we can construct a
class of bandpass updating schemes
that are also optimal for asymptotic convergence.
%
The optimal schedule of these bandpass schemes
is always given by an inverse-time formula.

@@

To show Eq. \eqref{eq:varY},
we first note that
the difference, $V_i - {\bar V}$,
differs from the actual deviation $x_i$ of bin $i$,
as defined in Eq. \eqref{eq:x_def},
only by a constant shift,
$\ln( p^*_i / p_i ) - \sum_{ j = 1 }^n \rho_j \, \ln( p^*_j / p_j )$.
%
This means that the two would share the same variance
under time averaging.
%
It follows that the corresponding linear combinations,
$Y_k$ and $y_k$, would also share the same variance.
%
But since the average of $y_k$ is zero,
we have
$$
\operatorname{var} Y_k
=
\operatorname{var} y_k
=
\left\langle y_k^2 \right\rangle.
$$
Finally, by using Eq. \eqref{eq:y2_eql},
we reach Eq. \eqref{eq:varY}.

@@

Next, we show that this bound represents
updating schemes with a set of very different eigenvalues.
%
First, observe that the summand of Eq. \eqref{eq:mass_func},
$\Gamma_k \, \left( \lambda_k e^{-\lambda_k \, Q} \right)^2$
is a function that peaks at $\lambda_k = 1/Q$,
so that the sum is dominated by the term at $k = k(Q)$ with
%
\begin{equation}
  \lambda_{k(Q)}
  \approx
  \frac 1 Q
  .
  \label{eq:kQ_def}
\end{equation}
%
Thus,
\begin{align}
  M(Q)
  &
  \gtrsim
  %\approx
  \sqrt{ \Gamma_{ k(Q) } } \,
  \lambda_{ k(Q) } \,
  e^{ -\lambda_{ k(Q) } \, Q }
\label{eq:MQ_lbound1}
  \\
  &
  \approx
  \frac{ \sqrt{ \Gamma_{ k(Q) } } }
       { Q }
  e^{ -\lambda_{ k(Q) } \, Q }
  .
\label{eq:MQ_lbound}
\end{align}

Now consider the $k$th error component, % of the asymptotic error,
\begin{equation}
  \left\langle
    y_k^2
  \right\rangle_A
  =
  \frac C T \,
  \Gamma_k \, \lambda_k^2
  \int_0^\infty
    \frac{ e^{-2 \, \lambda_k \, Q} }
         { M(Q) }
    \, dQ
  .
\notag
\end{equation}
%
Using Eq. \eqref{eq:MQ_lbound}, we have
$$
\begin{aligned}
  \left\langle
    y_k^2
  \right\rangle_A
  &
  \lesssim
  \frac C T \, \Gamma_k \, \lambda_k^2
  \int_0^\infty
    \frac{ Q } { \sqrt{ \Gamma_{ k(Q) } } }
    e^{-\left[ 2\,\lambda_k - \lambda_k(Q) \right] Q }
    \, dQ
  \\
  &
  \approx
  \frac C T \,
  \frac{ \Gamma_k } { \sqrt{ \Gamma_{ k(Q^*) } } }
  \left[
    \frac{ \lambda_k } { 2 \, \lambda_k - \lambda_{ k(Q^*) } }
  \right]^2
  ,
\end{aligned}
$$
where we have approximated $k(Q)$
by its value around the peak of the integrand
$$
  Q^*
  =
  \frac{ 1 } { 2 \, \lambda_k - \lambda_{ k(Q^*) } }
  .
$$
%
Comparing this to Eq. \eqref{eq:kQ_def},
we identify $k(Q^*) \approx k$, so
\begin{equation}
  \left\langle
    y_k^2
  \right\rangle_A
  \lesssim
  %\approx
  \frac C T
  \sqrt{ \Gamma_k }
  .
  \label{eq:y2_ubound}
\end{equation}
%
This result shows that with a set of well-separated eigenvalues,
the asymptotic error profile would be
independent of $\lambda_k$'s and
be flatter than the initial one from
Eq. \eqref{eq:y2_eql}
(cf. Fig. \ref{fig:xerr}
for a Gaussian example).
%
Note that we can turn Eq, \eqref{eq:y2_ubound}
to an inequality by choosing $k(Q) \to k$
for Eq. \eqref{eq:MQ_lbound1}.
%
Then by summing over $k$
and using Eq. \eqref{eq:mint_ubound},
we recover Eq. \eqref{eq:error_asym_ubound}.

For a finite-length simulation,
the integrals $\int_0^{q(T)} dQ$
can be extended to infinity
only if $q(T) \gg Q^*$.
%
In other words,
the above asymptotic limit fails around
$Q^* \approx q(T)$, or equivalently
around mode $k^* = k\bigl( q(T) \bigr)$,
where $\lambda_{ k^* } \approx 1/q(T)$.
%
\note{This case also exemplifies
  the sensitivity to the simulation length.
  %
  In this limit,
  $M(Q) \propto 1/Q$ by Eq. \eqref{eq:MQ_lbound},
  and $C \propto \ln\bigl[ q(T) / q_0 \bigr]$
  for some $q_0$,
  and we expect that with fixed $a_0$,
  $q(T)$ to increase almost linearly with $T$
  by Eq. \eqref{eq:opt_qT}.
  %
  So the normalized asymptotic error,
  $E_A \cdot T \propto \bigl( \ln[ q(T)/q_0 ] \bigr)^2 \sim (\ln T)^2$,
  is sensitive to the simulation length,
  and worse than the ideal case of a constant.
}

>>>>

First, we need to define the error in the bias potential.
%
This is done by removing two contributions from the bias potential.
%
The first contribution comes from the target value,
Eq. \eqref{eq:Vi_target}.
%
For this purpose,
we introduce a shifted bias potential
%
\begin{equation}
  v_i(t)
  \equiv
  V_i(t)
  -
  \ln \frac { p^*_i }
            { p_i }
  ,
  \label{eq:v_def}
\end{equation}
%
which should tend to a constant of $i$
upon convergence.
Note that the updates to $v_i$'s are
equivalent to the updates to $V_i$'s,
since the shift
remains a constant during the course of simulation,
%%
%In terms of $v_i$'s, Eq. \eqref{eq:pi_p_V}
%becomes
%%
%\begin{equation}
%  \pi_i
%  =
%  \frac{                p_i \, e^{-v_i} }
%       { \sum_{j = 1}^n p_j \, e^{-v_j} }
%  \propto
%  p_i \, e^{-v_i}.
%  \label{eq:pi_p_v}
%\end{equation}
%%
%According to Eq. \eqref{eq:pi_p_V1},

Second, the bias potential following Eq. \eqref{eq:wl_update}
generally grows over time.
%
Since a uniform increase of $V_i(t)$, or, equivalently, that of $v_i(t)$,
does not affect the
resulting distribution $\pi_i(t)$,
%
the real deviation is more readily defined
from the difference between $v_i(t)$ and a baseline value, $\bar v(t)$,
that represents the average growth\cite{
dama2014},
%
\begin{equation}
  x_i(t) \equiv v_i(t) - \bar v(t)
  %=
  %v_i - \sum_{j = 1}^n \rho_j \, v_j
  ,
\label{eq:x_def}
\end{equation}
%
where
%
\begin{equation}
  \bar v(t)
  = \sum_{i = 1}^n \rho_i \, v_i(t)
  ,
\label{eq:vbar_def}
\end{equation}
for some distribution,
$\pmb \rho = (\rho_1, \dots, \rho_n)$.
%
In this way,
the variable, $x_i(t)$,
represents the error of the bias potential,
and is expected to vanish on convergence.
%
We, therefore, define
the error of the bias potential as
%
\begin{equation}
  \Err
  =
  \sum_{i = 1}^n \rho_i \, \left\langle x_i^2(t) \right\rangle
  .
\label{eq:error_sum}
\end{equation}

@@

To proceed, we will
approximate the finite difference Eq. \eqref{eq:wl_update}
by a differential equation
%
\begin{equation}
  \dot V_i(t)
  \equiv
  \frac{ dV_i(t) } { dt }
  =
  \alpha(t) \, \frac{ h_i(t) } { p_i },
  \label{eq:vt_diffeq}
\end{equation}
%
with
%
$
h_i(t) = \delta_{i, i(t)}
$
%\begin{equation}
%  h_i(t) = \delta_{i, i(t)}
%  ,
%  \label{eq:h_def}
%\end{equation}
%
being the instantaneous histogram,
which is equal to $1.0$
for the bin $i(t)$ at time $t$
or zero otherwise.
%
The shifted bias potential, $v_i(t)$,
follows the same differential equation
as the shift is a constant.

Next, we split $h_i(t)$ into a deterministic averaged part
and a stochastic part, $\zeta_i(t)$,
%
\begin{equation}
  h_i(t) = \langle h_i(t) \rangle + \zeta_i(t).
  \label{eq:h_split}
\end{equation}
%
The deterministic part can be related
to the ``ensemble average'',
where the ensemble consists of many similar simulation copies
sharing the same schedule $\alpha(t)$
the same bias potential at time $t$.
%
The initial states and the stochastic forces
during the process may differ, however.

@@

%The average growth rate of $v(t)$ is given by
%\begin{equation}
%  \frac{ d \bar v (t) }
%       { d t }
%  =
%  \alpha(t)
%  \sum_{ r = 1 }^n \rho_r \, \frac{ h_r(t) } { p_r }
%  .
%\label{eq:dvbardt}
%\end{equation}
%

From Eqs.
\eqref{eq:v_def}-\eqref{eq:vbar_def},
\eqref{eq:rhox_sum}-\eqref{eq:h_ave},
%\eqref{eq:v_def},
%\eqref{eq:x_def},
%\eqref{eq:vbar_def},
%\eqref{eq:rhox_sum},
%\eqref{eq:vt_diffeq},
%\eqref{eq:h_split}
%\eqref{eq:h_ave},
%\eqref{eq:dvbardt},
we get a set of decoupled equations
%
\begin{equation}
  \dot x_i(t)
  =
  -\alpha(t) \, \left[ x_i(t) - \xi_i(t) \right]
  ,
\label{eq:dxdt_WL}
\end{equation}
where
\begin{equation}
  \xi_i(t)
  =
  \frac{ \zeta_i(t) }
       { p_i }
  -
  \sum_{ r = 1 }^n
    \rho_r
    \frac{ \zeta_r(t) }
         {     p_r    }
  .
\label{eq:xi_def}
\end{equation}
\note{The derivation:
$$
\begin{aligned}
  \dot x_i
  &\stackrel{\mathrm{Eq.\; \eqref{eq:x_def}}}
            {=\joinrel=\joinrel=\joinrel=\joinrel=}
  \dot v_i - \dot{ \bar v }
  \stackrel{\mathrm{Eqs.\; \eqref{eq:vt_diffeq}\; and\; \eqref{eq:vbar_def}}}
            {=\joinrel=\joinrel=\joinrel=\joinrel=\joinrel=\joinrel=\joinrel=\joinrel=\joinrel=\joinrel=}
  \alpha(t) \left(
    \frac{ h_i } { p_i }
    -
    \sum_{ r = 1 }^n
      \rho_r \frac{ h_r } { p_r }
  \right)
  \\
  &\stackrel{\mathrm{Eqs.\; \eqref{eq:h_split} and \; \eqref{eq:h_ave}}}
            {=\joinrel=\joinrel=\joinrel=\joinrel=}
  \alpha(t) \left(
    \frac{ \langle h_i \rangle } { p_i }
    -
    \sum_{ r = 1 }^n
      \rho_r \frac{ \langle h_r \rangle } { p_r }
    +
    \xi_i(t)
  \right)
  \\
  &\stackrel{\mathrm{Eqs.\; \eqref{eq:rhox_sum} \;and\; \eqref{eq:h_ave}}}
            {=\joinrel=\joinrel=\joinrel=\joinrel=\joinrel=\joinrel=\joinrel=\joinrel=\joinrel=\joinrel=\joinrel=}
  \alpha(t) \left[ -x_i(t) + \xi_i(t) \right].
\end{aligned}
$$
}

The total error defined in Eq. \eqref{eq:error_sum}
now depends implicitly on the function $\alpha(t)$
via $x_i(t)$.
%
Our aim is to find the $\alpha(t)$
that minimizes the error
at the end of the simulation period $T$.
%
It suffices to consider
a one-variable problem that
minimizes each individual
$\left\langle x_i^2(T) \right\rangle$.

@@

The error function defined in Eq. \eqref{eq:error_sum}
can be rewritten in terms of the eigenmodes of $\mathbf w$
\begin{align}
  \Err
  \equiv
  \sum_{i = 1}^n \rho_i \,
    \left\langle x_i^2 \right\rangle
  =
  \sum_{i, j, k=1}^n \phi_{ki} \, \phi_{kj} \,
    \left\langle x_i \, x_j \right\rangle
  =
  \sum_{k = 1}^n
    \left\langle
      y_k^2
    \right\rangle
  ,
\notag
\end{align}
%
where we have used Eq. \eqref{eq:eig_orthonormal_cols}.
%
From Eq. \eqref{eq:eigenmode1},
the first eigenmode,
$y_1 = \sum_{j=1}^n \rho_j \, x_j$,
vanishes identically by Eq. \eqref{eq:rhox_sum}.
%
Thus, we can start the last sum %in Eq. \eqref{eq:y2_sum}
from $k = 2$,
\begin{equation}
  \Err
  =
  \sum_{k = 2}^n
    \left\langle
    y_k^2(T)
    \right\rangle
  .
\label{eq:y2_sum}
\end{equation}

@@

We have also shown that
the single-bin scheme is optimal
in the long time limit
without a priori assumptions
about the smoothness of the PMF.
%
However, if certain eigenmodes for local noises
are expected to vanish in the PMF,
we can construct a
class of bandpass updating schemes
that are also optimal for asymptotic convergence.
%
The optimal schedule of these bandpass schemes
is always given by an inverse-time formula.

@@

To show Eq. \eqref{eq:varY},
we first note that
the difference, $V_i - {\bar V}$,
differs from the actual deviation, $x_i$,
defined in Eq. \eqref{eq:x_def},
only by a constant shift,
$\ln( p^*_i / p_i ) - \sum_{ j = 1 }^n \rho_j \, \ln( p^*_j / p_j )$.
%
Thus, the two share the same variance
under time averaging.
%
It follows that the corresponding linear combinations,
$Y_k$ and $y_k$, would also share the same variance.
%
But since the average of $y_k$ is zero,
we have
$$
\operatorname{var} Y_k
=
\operatorname{var} y_k
=
\left\langle y_k^2 \right\rangle.
$$
Finally, by using Eq. \eqref{eq:y2_eql},
we reach Eq. \eqref{eq:varY}.

@@

Next, we will show that this bound represents
updating schemes with a set of very different eigenvalues.
%
%First, observe that the summand of Eq. \eqref{eq:mass_func},
%$\Gamma_k \, \left( \lambda_k e^{-\lambda_k \, Q} \right)^2$
%is a function that peaks at $\lambda_k = 1/Q$,
%so that the sum is dominated by the term at $k = k(Q)$ with
%%
%\begin{equation}
%  \lambda_{k(Q)}
%  \approx
%  \frac 1 Q
%  .
%  \label{eq:kQ_def}
%\end{equation}
%
%Thus,
First, we have for any mode $k^*$ $(k^* \ge 2)$,
%
\begin{align}
  M(Q)
  &
  \ge
  %\approx
  \sqrt{ \Gamma_{ k^* } } \,
  \lambda_{ k^* } \,
  e^{ -\lambda_{ k^* } \, Q }
  ,
\label{eq:MQ_lbound1}
\end{align}
%
which would approach an equality if the $k^*$ mode
dominates the mass function, $M(Q)$,
defined in Eq. \eqref{eq:mass_func}.
%
Then, for the $k$th component of the asymptotic error,
we have from Eq. \eqref{eq:error_asym_Lagrangian}
%
\begin{align}
  \left\langle
  y_k^2(T)
  \right\rangle_A
  &=
  \int_0^T
  \Gamma_k \, \lambda_k^2 \, e^{2 \, \lambda_k \, [q(t) - q(T)] }
  \, {\dot q}^2(t) \, dt
  \notag
  \\
  &=
  \frac{ C } { T }
  \, \Gamma_k \, \lambda_k^2
  \int_0^{q(T)}
  \frac{ e^{-2 \, \lambda_k \, Q } } { M(Q) }
  \,  \, dQ
  \notag
  \\
  &\le
  \frac C T \,
  \frac{ \Gamma_k \, \lambda_k^2 } { \sqrt{ \Gamma_{k^*} } \, \lambda_{k^*} }
  \int_0^{q(T)}
    e^{-(2 \, \lambda_k - \lambda_{k^*} ) \, Q}
    \, dQ
  \notag
  \\
  &\le
  \frac C T \,
  \cdot
  \frac{ \Gamma_k }
  { \sqrt{ \Gamma_{k^*} } }
  \cdot
  \frac{ \lambda_k^2 }
  { \lambda_{k^*} \, (2 \, \lambda_k -\lambda_{k^*}) }
  ,
  \label{eq:y2_ubound_k}
\end{align}
%
where we have
used Eqs. \eqref{eq:Lagrangian_const} and \eqref{eq:MQ_lbound1}
on the second and third lines, respectively,
and extended the upper limit to infinity on the last line.
%
For a set of well-separated eigenvalues,
$\lambda_k$ depends on the mode index $k$
much more critically than $\Gamma_k$.
%
Thus, the tightest upper bound
can be approximately obtained from minimizing the last factor
on the right-hand side of Eq. \eqref{eq:y2_ubound_k},
and the minimal is reached at $k^* = k$,
when $\lambda_{k^*} = \lambda_k$.
%
So
%
\begin{equation}
  \left\langle
  y_k^2(T)
  \right\rangle_A
  \le
  %\approx
  \frac C T
  \sqrt{ \Gamma_k }
  .
  \label{eq:y2_ubound}
\end{equation}
%
By summing over $k$
and using Eq. \eqref{eq:mint_ubound},
we recover Eq. \eqref{eq:error_asym_ubound}.

This result shows that with a set of well-separated eigenvalues,
the asymptotic error profile would be
independent of $\lambda_k$'s and
be flatter than the initial one from
Eq. \eqref{eq:y2_eql}
(cf. Fig. \ref{fig:xerr}
for a Gaussian example).
%
Note, however, that to approach this upper bound
the simulation has to be long enough
such that the upper limit of the integral
can be extended from $q(T)$ to infinity
in the last step of Eq. \eqref{eq:y2_ubound_k}.
%
As a result,
the flatter profile from Eq. \eqref{eq:y2_ubound}
would stop at mode $k$,
at which $\lambda_k \approx 1/q(T)$.
%In other words,
%the above asymptotic limit fails around
%$Q^* \approx q(T)$, or equivalently
%around mode $k^* = k\bigl( q(T) \bigr)$,
%where $\lambda_{ k^* } \approx 1/q(T)$.
%%
%\note{This case also exemplifies
%  the sensitivity to the simulation length.
%  %
%  In this limit,
%  $M(Q) \propto 1/Q$ by Eq. \eqref{eq:MQ_lbound},
%  and $C \propto \ln\bigl[ q(T) / q_0 \bigr]$
%  for some $q_0$,
%  and we expect that with fixed $a_0$,
%  $q(T)$ to increase almost linearly with $T$
%  by Eq. \eqref{eq:opt_qT}.
%  %
%  So the normalized asymptotic error,
%  $E_A \cdot T \propto \bigl( \ln[ q(T)/q_0 ] \bigr)^2 \sim (\ln T)^2$,
%  is sensitive to the simulation length,
%  and worse than the ideal case of a constant.
%}
